---
title: "Module tests for plot_OSLcurve()"
author: "Dirk Mittelstraß"
date: "2025-08-31"
format:
  html:
    code-fold: true 
    code-summary: "Show code"
    toc: true
    toc-location: right
    toc-expand: true
    self-contained: true
editor: visual
execute:
  output: true
  warning: true
---

This document collects scripts for developing, bug-fixing and pre-release validating a specific function.

**Found issues as for 2025-08-31:**

-   If `show.intervals = TRUE` there is an unimportant "scale is overwritten" warning

-   There are too many warnings when values are omitted due to log scaling. One warning would be enough

-   If a simulated curve is used as input but no component given, the warnings are not helpful

**ToDo:**

-   Test if function works with background component

-   More tests for RLum input objects

## Test sourcing

Is the function successfully sourced?

```{r sourcing}

library(OSLdecomposition)
library(ggplot2)
library(Luminescence)

source("../R/plot_OSLcurve.R")
source("../R/plot_MultiExponential.R") # core function

```

## Define test cases

```{r cases}

cases <- list()
add_component_case <- 
  function(cases, lambda, n, 
           channel_width = 0.1, 
           channel_number = 400, 
           noise = 0, 
           background = 0,
           reiterate = FALSE,
           title = "Unspecified test case"){
    
    components <- data.frame(name = paste("Component", 1:length(n)),
                             lambda = lambda, 
                             n = n)
    curve <- simulate_OSLcomponents(components, 
                                    channel.width = channel_width,
                                    channel.number = channel_number,
                                    simulate.curve = TRUE, 
                                    add.poisson.noise = TRUE,
                                    add.gaussian.noise = noise,
                                    add.background = background)
    
    # Re-iterate components to add some uncertainty
    if (reiterate) {
          components <- decompose_OSLcurve(curve, 
                                     components, 
                                     verbose = FALSE)
    }
  
    case <- list(components = components,
                 curve = curve,
                 title = title)
    
    cases <- c(cases, list(case))
    return(cases)
  }

cases <- add_component_case(cases,
                   lambda = c(1, 0.2, 0.02),
                   n = c(1000, 1000, 10000),
                   reiterate = TRUE,
                   title = "Typical quartz OSL curve")

cases <- add_component_case(cases,
                   lambda = c(2, 0.5, 0.2, 0.001),
                   n = c(200, -500, 2000, 100000),
                   channel_width = 0.875,
                   channel_number = 42,
                   title = "Atypical low-res measurement with stretched medium component")

cases <- add_component_case(cases,
                   lambda = c(0.1, 0.02),
                   n = c(-10e4, 10e5),
                   channel_width = 0.2,
                   channel_number = 200,
                   title = "Increasing signal")

cases <- add_component_case(cases,
                   lambda = c(1, 0.01),
                   n = c(10, 100),
                   noise = 10,
                   reiterate = TRUE,
                   title = "Background-corrected noise")

cases <- add_component_case(cases,
                   lambda = c(2, 0.5, 0.03, 0.001),
                   n = c(30e5, 2e6, -5e6, 5e8),
                   channel_width = 0.01,
                   channel_number = 5000,
                   title = "Sharp drop then complex, with many channels")


```

## Test standard plotting

### Default settings

```{r default}

for (case in cases) {
  plot_OSLcurve(case$curve, case$components,
                       title = case$title)
  cat("\n")
}
```

### Changed on/off settings

```{r table_settings}

for (case in cases) {
  plot_OSLcurve(case$curve, case$components,
                show.legend = FALSE,
                show.intervals = TRUE,
                show.crosssec = TRUE,
                show.initial = TRUE)
  cat("\n")
}

```

## Test other display modes

Vary the argument "display". Vary also the ggplot2 themes

```{r display_settings}

modes <- c("lin","compare_lin","log","compare_log","loglog","LM","res","tab","raw")
themes <- list(ggplot2::theme_minimal(),
            ggplot2::theme_dark(),
            ggplot2::theme_classic(),
            ggplot2::theme_void(),
            ggplot2::theme_gray(),
            ggplot2::theme_linedraw(),
            ggplot2::theme_bw())

for (display_mode in modes) {
  
  case <- sample(cases, size = 1)[[1]]  # Random case
  cat("Display mode:", display_mode, "with case", case$title, "\n")
  

  
    tryCatch({
      plot_OSLcurve(case$curve, case$components,
                display = display_mode,
                show.legend = TRUE,
                theme.set = sample(themes, size = 1)[[1]],
                title = case$title)
    
    },error = function(e) {message("FOUND ERROR", conditionMessage(e))})
  
  cat("\n")
}

# Repeat test but with 

for (display_mode in modes) {
  
  case <- sample(cases, size = 1)[[1]]  # Random case
  cat("Display mode:", display_mode, "with case", case$title, "\n")
  
  tryCatch({
    plot_OSLcurve(case$curve, case$components,
                display = display_mode,
                show.legend = FALSE,
                show.intervals = TRUE,
                theme.set = sample(themes, size = 1)[[1]])
    
    },error = function(e) {message("FOUND ERROR", conditionMessage(e))})
  
  cat("\n")
}
```

## Display fit_OSLcurve output

```{r fit_OSLcurve}
# Create a simple curve with just one component
curve <- data.frame(
  X = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12),
  Y = c(377, 244, 163, 93, 59, 28, 17, 13, 10, 8, 9, 5))
# Perform fitting
components <- fit_OSLcurve(curve, F.threshold = 3)

# Display results
plot_OSLcurve(curve, components)

```

## Use RLum objects as input

```{r rlum_input}

# Load example data set
data_path <- system.file("examples", 
                         "FB_10Gy_SAR.bin", 
                         package = "OSLdecomposition")
data_set <- Luminescence::read_BIN2R(data_path, 
                                     fastForward = TRUE,
                                     verbose = FALSE)

# Separate components
data_set_decomposed <- RLum.OSL_decomposition(data_set, 
                                              decay_rates = c(0.8, 0.05))

cat("\nCase 1: Use raw RLum.Data.Curve as input\n")
plot_OSLcurve(data_set[[4]]@records[[1]])

cat("\nCase 2: Use decomposed RLum.Data.Curve as input\n")
plot_OSLcurve(data_set_decomposed[[4]]@records[[1]])

cat("\nCase 3: Use data matrix as input\n")
plot_OSLcurve(data_set_decomposed[[4]]@records[[1]]@data)

cat("\nCase 4: Use invalid RLum object as input\n")
tryCatch({plot_OSLcurve(data_set[[4]])}, 
         error = function(e) {
           message("TEST ERROR ", conditionMessage(e))
})

```

## Wrong/missing input

```{r wrong_input}

case <- sample(cases, size = 1)[[1]]  # Random case

cat("\nCase 1: No component argument\n")
tryCatch({plot_OSLcurve(case$curve)}, 
         error = function(e) {
           message("TEST ERROR ", conditionMessage(e))
})

cat("\nCase 2: No curve argument\n")
tryCatch({plot_OSLcurve(component = case$component)}, 
         error = function(e) {
           message("TEST ERROR ", conditionMessage(e))
})

cat("\nCase 3: No arguments at all\n")
tryCatch({plot_OSLcurve()}, 
         error = function(e) {
           message("TEST ERROR ", conditionMessage(e))
})

cat("\nCase 4: Wrong argument\n")
tryCatch({plot_OSLcurve(case$component, case$curve)}, 
         error = function(e) {
           message("TEST ERROR ", conditionMessage(e))
})
```

## Don't show plot

```{r hide_plot}

case <- sample(cases, size = 1)[[1]]  # Random case
plot_obj <- plot_OSLcurve(case$curve, case$components, hide.plot = TRUE)

# Try to display by itself
plot_obj

# Try to display with print()
print(plot_obj)

# Try to display with gridExtra function
gridExtra::grid.arrange(plot_obj)

```

## Save plot as file it

The first plot is saved as PNG file and shown, the second plot is saved as PDF file and not shown. Both are saved in the working directory (`r getwd()`). Please check and delete both files after executing this test.

```{r save_plot}

file_name = paste0(getwd(), "/plot_OSLcurve_test.png")
cat("\nSave file:", file_name, "\n")
case <- sample(cases, size = 1)[[1]]  # Random case

plot_OSLcurve(case$curve, case$components,
              hide.plot = FALSE,
              filename = file_name)
cat("\n")

file_name = "plot_OSLcurve_test.pdf"
cat("\nSave file:", file_name, "\n")
case <- sample(cases, size = 1)[[1]]  # Random case

plot_OSLcurve(case$curve, case$components,
              hide.plot = TRUE,
              filename = file_name)
cat("\n")

```
