---
title: "Module tests for sum_OSLcurves()"
author: "Dirk Mittelstraß"
date: "2026-03-12"
format:
  html:
    toc: true
    toc-location: right
    toc-expand: true
    self-contained: true
editor: visual
execute:
  fig-width: 10
  fig-height: 5
  output: true
  warning: true
---

This document collects scripts for developing, bug-fixing and pre-release validating a specific function.

## Sourcing

Is the function successfully sourced?

```{r sourcing}

library(OSLdecomposition)
source("../R/sum_OSLcurves.R")

```

## Load data sets

Data sets for testing

```{r}

library(Luminescence)

# 'FB_10Gy' is a dose recovery test with the Fontainebleau quartz
# measured in a lexsyg research with green LED stimulation
data_path <- system.file("examples", 
                         "FB_10Gy_SAR.bin", 
                         package = "OSLdecomposition")
data_FB <- read_BIN2R(data_path, fastForward = TRUE, verbose = FALSE)

# Load another example data from the Luminescence package
data(ExampleData.BINfileData, envir = environment())
data_Lum <- Risoe.BINfileData2RLum.Analysis(CWOSL.SAR.Data)
```

## Input object types

```{r test_calc}

# Give average CW-OSL curve of back
m_curve <- sum_OSLcurves(data_FB, output.plot = TRUE)

# Does the red average curve match the actually returned curve?
plot(m_curve, type = 'l')

# Second allowed object type
record_list <- data_FB[[3]]@records
m_curve <- sum_OSLcurves(record_list, output.plot = TRUE)


# Third allowed object type
df_list <- list(as.data.frame(record_list[[1]]@data),
                as.data.frame(record_list[[2]]@data),
                as.data.frame(record_list[[5]]@data))

m_curve <- sum_OSLcurves(df_list, 
                         output.plot = TRUE, 
                         plot.first = TRUE,
                         title = "auto")


# This is not allowed and should return NULL
sum_OSLcurves(c(1:13), output.plot = TRUE)
```

## Selecting record types

Test extracting different types of records and again plotting

```{r test_plots}
# Regular OSL measurements
table <-sum_OSLcurves(data_Lum, record_type = "OSL", output.plot = TRUE)

# Should return average of 24 post-OSL IRSL tests
sum_OSLcurves(data_Lum, record_type = "IRSL", output.plot = TRUE,
              title = "Average post-OSL IRSL tests")

# Should return average of preheat TL curves
sum_OSLcurves(data_Lum, record_type = "TL", output.plot = TRUE, 
              title = "Average of preheat TL curves")

# Should return NULL
sum_OSLcurves(data_Lum, record_type = "TROSL", output.plot = TRUE)
```

## Selecting subsets

Calculate just from a subset, needed for `RLum.OSL_correction`

```{r test_selection}

# Should return background curve
sum_OSLcurves(data_FB, record_type = "OSL", 
              selection = 11,
              output.plot = TRUE,
              title = "FB quartz: background curve")

# Should return a subset
sum_OSLcurves(data_Lum, record_type = "OSL", 
              selection = 11:13,
              output.plot = TRUE,
              title = "Only three aliquots")

# Should return a error message
result <- tryCatch({

  sum_OSLcurves(data_FB, selection = 11:17)
  
}, error = function(w) {
  cat("ERROR:", conditionMessage(w), "\n")
})
```
