---
title: "Module tests for decompose_OSLcurve()"
author: "Dirk Mittelstraß"
date: "2024-09-25"
format:
  html:
    toc: true
    toc-location: right
    toc-expand: true
    self-contained: true
editor: visual
execute:
  output: true
  warning: true
---

This document collects scripts for developing, bug-fixing and pre-release validating a specific function.

## Sourcing

Is the function successfully sourced?

```{r sourcing}

library(OSLdecomposition)
source("../R/decompose_OSLcurve.R")

# For measuring execution time
library(tictoc)

# For printing tables
library(knitr)
```

## Basic algorithm test

Create test curve

```{r}

# Some common quartz CW-OSL measurement with a PMT
sim_components <- data.frame(name = c("Fast", "Medium", "Slow"),
                             lambda = c(1.5, 0.2, 0.01), 
                             n = c(1000, 1000, 20000),
                             n.error = NA)

# Simulate the curve, add shot noise, detector noise and detector background
curve <- simulate_OSLcomponents(sim_components, 
                                simulate.curve = TRUE, 
                                add.poisson.noise = TRUE,
                                add.gaussian.noise = 2,
                                add.background = 10)

# How does the curve look like?
plot(x = curve$time, y = curve$signal, type = "l")

```

Cycle algorithms, without background correction

```{r}

results <- data.frame(NULL)

for (arg_algorithm in c("det", 
                        "nls", 
                        "det+nls")) {
  for (arg_error_est in c("empiric", 
                          "poisson", 
                          "only.bin.RSS", 
                          "none", 
                          2)) {
    
    case_title <- paste("Algorithm =", arg_algorithm,
                        "| Error est. =", arg_error_est)
    cat("\n", case_title, "\n")
    
    # Function call with return
    components <- decompose_OSLcurve(curve,
                                     sim_components,
                                     background.fitting = FALSE,
                                     algorithm = arg_algorithm,
                                     error.estimation = arg_error_est,
                                     verbose = TRUE)
    
    
    # Call the function now 1000 times to test performance and see if errors occur
    tic("Execution time")
    for (i in 1:1000) {
      suppressWarnings({
        decompose_OSLcurve(curve,
                           sim_components,
                           background.fitting = FALSE,
                           algorithm = arg_algorithm,
                           error.estimation = arg_error_est,
                           verbose = FALSE)
      })
    }
    tictoc <- toc()
    elapsed <- as.numeric((tictoc$toc - tictoc$tic)/1000)
    
    result <- data.frame(test = case_title,
                         t(components$n),
                         t(components$n.error),
                         time = elapsed)
    
    results <- rbind(results, result)
    
  }
}

kable(results)
```

### With background correction

Now do the same, **with** background correction.

```{r}

# Background fittings needs pre-defined integration intervals
sim_components <- optimise_OSLintervals(sim_components, curve, 
                                        background.component = TRUE)

results <- data.frame(NULL)

for (arg_algorithm in c("det", 
                        "nls", 
                        "det+nls")) {
  for (arg_error_est in c("empiric", 
                          "poisson", 
                          "only.bin.RSS", 
                          "none", 
                          2)) {
    
    case_title <- paste("Algorithm =", arg_algorithm,
                        "| Error est. =", arg_error_est)
    cat("\n", case_title, "\n")
    
    # Function call with return
    components <- decompose_OSLcurve(curve,
                                     sim_components,
                                     background.fitting = TRUE,
                                     algorithm = arg_algorithm,
                                     error.estimation = arg_error_est,
                                     verbose = TRUE)
    
    
    # Call the function now 1000 times to test performance and see if errors occur
    tic("Execution time")
    for (i in 1:1000) {
      suppressWarnings({
        decompose_OSLcurve(curve,
                           sim_components,
                           background.fitting = TRUE,
                           algorithm = arg_algorithm,
                           error.estimation = arg_error_est,
                           verbose = FALSE)
      })
    }
    tictoc <- toc()
    elapsed <- as.numeric((tictoc$toc - tictoc$tic)/1000)
    
    result <- data.frame(test = case_title,
                         t(components$n),
                         t(components$n.error),
                         time = elapsed)
    
    results <- rbind(results, result)
  }
}

kable(results)
```

## Input checks

Try out a variety of possible cases for the lambda input

```{r}

component_inputs <- list(as_usual = sim_components,
                        lambda_only = sim_components$lambda,
                        named_vector = c(Schnell = 1,
                                         Mittel = 0.1,
                                         Langsam = 0.01),
                        single_lambda = 0.5,
                        many_lambda = c(2,1,0.5,0.2,0.1,0.05, 0.01))

for (i in 1:length(component_inputs)) {
  
  cat("\nTest case:", names(component_inputs)[i], "\n")
   decompose_OSLcurve(curve, component_inputs[[i]])
}

```
