---
title: "Step1_report"
author: "Dirk Mittelstrass"
date: "2020/04/03"
output:
  html_document:
    df_print: paged
params:
  data_set: !r list(NULL)
---

```{r setup, include=FALSE}

library(numOSL)
library(knitr)
library(ggplot2)
library(gridExtra)
library(ggpubr)

# fig.width=6,
knitr::opts_chunk$set(fig.width=7,
                      fig.asp=.4,
                      results = "asis",
                      warning=FALSE,
                      message=FALSE,
                      error=FALSE,
                      echo=FALSE,
                      cache=FALSE)

data_set <- params$data_set


```

## Test

... fill intro here ...

```{r test}

length(data_set)
```

The subsequent diagrams are structured the following way:

* Upper left: Global mean curve (grey), fit model curve (black) and component signals
* Upper right: Same as log-log diagram
* Lower left: Residual curve between fit and global mean curve
* Lower right: Result table with estimated type of component names (colored)


```{r step1_display_cases, fig.asp=.6}

# display fittings for all K
for (i in 1:nrow(C.list$F.test)) {
  
  plot_OSLcurve(global_curve, 
                C.list$components.list[[i]], 
                title = NULL,
                algorithm = " ")

 # cat(paste0("<pre>**Figure ", n.figure,":** "))
  cat(paste0("<pre>*Figure ", n.figure,": Global mean curve fit with K = ", i, 
             " components* <pre>  ---------  <pre>"))
n.figure <- n.figure + 1

#cat("")
}

```

\pagebreak
**F-test**  
But which of these fittings gives back a sufficient model of the global mean curve, without over-fitting it? We solve this by comparing the residual square sum (*RSS*) of each fitting with the *RSS* value of the previous fitting. Bluszcz & Adamiec (2006) propose to use a *F*-test:  

$$F_K = \frac{(RSS_{K-1} - RSS_K)/2}{RSS_K(N - 2K)} $$


If *F~K~* falls below the preset threshold value of *F~threshold~* = `r F_threshold`, the new fitting model with *K* components is apparently not significantly better than the *K* - 1 model.


```{r F-test_table}
# print F-test table
kable(C.list$F.test.print, 
      escape = TRUE, align = "c",
      caption = "Table 1: Decay constants and fit quality parameters for multi-exponentional decay fitting with *K* components")
```

The fitting with *K* = `r K.selected` components is found to be the best suiting model to describe the given sample. Signal components with not-first-order kinetics, however, can lead to over-fitting. It is recommended to take the results of the *K* = `r K.selected - 1` fitting model also into consideration.  

If the sample was measured with a stimulation light wavelength of about 470 nm and a stimulation light intensity of `r stimulation_intensity` mW cm^-2^ as presetted, the photoionisation cross-sections of the components can be calculated. These can be compared with the quartz LM-OSL findings given in literature.

```{r step1_crosssections, fig.asp = 0.32}

#cross.plot.height <- 0
#if ((stimulation_wavelength >= 465) && (stimulation_wavelength <= 480)) cross.plot.height <- 3
#cross.plot.height <- (cross.plot.height + 2 + K.selected) * 0.01
#knitr::opts_current$set(fig.asp = cross.plot.height)

plot_PhotoCrosssections(C.list,
                        stimulation.intensity = stimulation_intensity,
                        stimulation.wavelength = stimulation_wavelength)

  cat(paste0("<pre>*Figure ", n.figure,": Comparison of decay constants between fitting cases and comparison with reference values. Red square: Best fit*"))
n.figure <- n.figure + 1

```
