---
title: "Step 2: Algebraic decomposition of all CW-OSL curves"
output:
  html_document:
    theme: journal
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: false
    df_print: paged
params:
  dec_data: !r list(NULL)
  data_set: !r list(NULL)
  object_name: !r c("")
---
Report of the [Mittelstraß *et al.* (2020)](#disclaimer) Step 2 analysis of the data set **`r object_name`** performed with the function `RLum.OSL_decomposition()` from the **R** package `OSLdecomposition`  `r packageVersion("OSLdecomposition")` executed at `r Sys.time()`.

***

```{r step2_setup, include=FALSE}

# Last changed: 2020-05-07

library(numOSL)
library(knitr)
library(ggplot2)
library(gridExtra)
#library(ggpubr)
library(kableExtra)

knitr::opts_chunk$set(fig.width=7,
                      fig.asp=.6,
                      results = "asis",
                      warning=FALSE,
                      message=FALSE,
                      error=FALSE,
                      echo=FALSE,
                      cache=FALSE)
## set ggplot design 
# (just for directly plotted diagrams. sub-function plot are not affected)
theme_set(theme_minimal())
graph_colors <- c("red3","green3","blue3","darkorchid","gold","brown","pink")

# numerate figures and tables
n.figure <- 1
n.table <- 1

# Find out the overall number of 'record_type' curves and the indicies of the OSL curves
# ToDo: This is ugly code. Refactorize, for example with grep()
record_index <- list()
n.curves <- 0

for (j in 1:length(data_set)) {
  record_index[[j]] <- c(NA)
  n.aliquot <- 0
  for (i in 1:length(data_set[[j]]@records)) {
    
   if (data_set[[j]]@records[[i]]@recordType == record_type) {
     n.curves <- n.curves + 1
     n.aliquot <- n.aliquot + 1
     # Collect index of OSL records
     record_index[[j]][n.aliquot] <- i
   }
  }
}

K <- nrow(dec_data$decompositon.input)
record_type <- dec_data$parameters$record_type
background_fitting <- dec_data$parameters$background_fitting
input_table <- dec_data$decompositon.input
```


## Basics 
<font color="black">

If we assume `r K` independent [exponential decays](https://en.wikipedia.org/wiki/Exponential_decay) to describe `r record_type` signal curves, then each single `r record_type` measurement can be approximated by:

```{r step2_draw_formula}
append_B <- ""
if (background_fitting) append_B <- "+B"

# Draw basic formula
cat("$$I(t) = ", paste(paste0("n_",1:K, "e^{\\lambda_",1:K,"t}"), collapse = "+"), append_B, "$$")
```


We assume that the decay rates $\lambda$ are similar in all measurements of the data set and define them as global constants. These global decay constans were already determined in Step 1 or preset by the user:

```{r step2_decay_rate_list}
for (k in 1:K) {
  cat(paste0("* ", input_table$name[k]," $\\lambda_", k,"=", round(input_table$lambda[k], digits = 4),"$ $s^{-1}$\n"))
}
```

To solve equation (1) for each measurement, we divide each signal curve into `r K` bins. The (known) signal value of each bin can be approximated  Each signal bin is then dexcribed by 
... add signal bin info here ...

... and fundamental approach here ...

... and refer to the paper or master thesis here ...


## Aliquot results

The processed data set consists of `r length(data_set)` `RLum.Analysis` objects. Each `RLum.Analysis` objects represents one measurement sequence and therefore the measurement of one aliquot or one grain in case of single grain data.

```{r step2_create_tab_data} 
### This chunk rearranges the data for the dynamic tabs, no direct output is produced

intensity_tables <- list()
barplots <- list()
barplot_labels <- list()

for (j in 1:length(data_set)) {
  result <- dec_data$results[dec_data$results$list.index == j,]
   if(result$IRR_TIME[1] == 0) result$IRR_TIME[1] <- "?"
  
  aliquot_table <- subset(result, select = c(record.index, IRR_TIME))
  
  plot_table <- data.frame(NULL)

  colnames(aliquot_table) <- c("record index", "irradiation time (s)")
  
  for (k in 1:K) {
    intensities <- result[,colnames(result) == paste0("n.", k)]
    intensities_reduced <- intensities - result[,colnames(result) == paste0("n.residual.", k)]
    sigmas <- result[,colnames(result) == paste0("n.error.", k)]
    
    # Append current component to Table 1 table
    new_col <- data.frame(row = paste0(round(intensities), " ± ", round(sigmas)))
    colnames(new_col) <- paste0(input_table$name[k],"\n$n_", k, " ± \\sigma_", k,#"$")
                                "$ (", input_table$name[k] ,")")
    aliquot_table <- cbind(aliquot_table, new_col)
    
    # Append current component to Figure 2 plot data
    plot_table <- rbind(plot_table, data.frame(x_axis = 1:length(intensities_reduced),
                                               y_axis = intensities_reduced,
                                               component = input_table$name[k]))
  }
  
  intensity_tables[[j]] <- aliquot_table
  barplots[[j]] <- plot_table
  barplot_labels[[j]] <- data.frame(breaks = 1:length(intensities_reduced),
                                    labels = paste0("#", 1:length(intensities_reduced), 
                                                    "\n", result$IRR_TIME, " s"))
  # add record@info[["IRR_UNIT"]]
}
```

### Intensitiy tables {.tabset .tabset-pills}

The decompositon process calculates the signal component intensities $n$ and their standard deviations $\sigma$. The intensity values $n$ equal the areas under the component curves (including the remaining virtual signal beyond the stimulation end) and have the same unit as the detector signal (usually *counts* or *mA*). Be aware that the uncertainty values $\sigma$ account just for the signal noise. Errors in the decay rate values or the decay kinetics may lead to intensity value $n$ discrepancies without increasing the uncertainty value $\sigma$ enough to sufficiently account for them.


```{r step2_create_result_tabs1} 
# Define dynamic Rmarkdown code, structured in tabs
dynamic_code <- c("#### #{{j}}\n",
                  
# Print the full intensity table
    "<i><b>Table 1.`r {{j}}`:</b> Signal component intensities $n$ and their standard deviation $\\sigma$ calculated by the decomposition process.</i><br>",
    "```{r tab_{{j}}_result_table}\n",
   # "kable_table <- intensity_tables[[{{j}}]]\n",
    "kable(intensity_tables[[{{j}}]], escape = TRUE, row.names = FALSE, align = \"c\") %>%\n",
    "kable_styling(bootstrap_options = c(\"striped\", \"condensed\"), fixed_thead = TRUE)",
    "```\n",
    "<br><br>\n")

tabs <- lapply(as.list(1:length(data_set)),
                function(j) knitr::knit_expand(text = dynamic_code))

# Now knit the dynamic code. This has to be in a seperate chunk
```
`r knitr::knit(text = unlist(tabs))`
<br>
<br>

### Intensity plots {.tabset .tabset-pills}
```{r step2_create_result_tabs2} 
dynamic_code <- c("#### #{{j}}\n",
                  
    # Draw the intensity barplot
    "```{r tab_{{j}}_result_barplot}\n",
    "ggplot(barplots[[{{j}}]], aes(fill=component, y=y_axis, x=x_axis)) + ",
    "geom_bar(position=\"dodge\", stat=\"identity\") + ",
    "scale_x_continuous(name = paste0(record_type, \" measurement no. and received dose\"),",
                     "breaks = barplot_labels[[{{j}}]]$breaks,",
                     "labels = barplot_labels[[{{j}}]]$labels) +",
    "scale_y_continuous(name = \"component signal\") +", #, labels = scientific
    "scale_fill_manual(values = graph_colors)",
    "```\n",
    "<i><b>Figure 1.`r {{j}}`:</b> `r record_type` component intensities of all measurements ",
    "of measurement sequence `r {{j}}`. Contrary to table 1, not the values of parameter $n$ ",
    "are displayed, instead the remaining signal after the stimulation end is subtracted from $n$. ",
    "This doesn't affect the signal behaviours but reduces values of slow decaying components ",
    "for better comparibility.",
    "   </i><br><br>\n")

tabs <- lapply(as.list(1:length(data_set)),
                function(j) knitr::knit_expand(text = dynamic_code))
```
`r knitr::knit(text = unlist(tabs))`

### Decay models {.tabset .tabset-fade .tabset-pills}

With all parameters of our fundamental equation now known (see first paragraph), the signal decay can now be modelled. If the scattering of the measured signal around the model curve is displayed by the residual plot (lower plots in figure 2). Is the residual plot dominated by random noise, the model can be expected as valid. Systematic deviations in parts of the plot indicate at least one of the following systematic errors:

* Incorrect decay constants 
* Not-first-order decay kinetics
* Insufficient number of signal components (under-fitting)
* Unstable light stimulation
* Inadequate or missing background correction

The plots in figure 2 allow the evaluation of the model quality for the first two measurements of each measurement sequence.

```{r step2_create_result_tabs3, eval = TRUE} 
# Define dynamic Rmarkdown code
dynamic_code <- c(
  
    # Tab header
    "#### #{{j}}\n",
    
    # Draw the OSL plots of the first two records side by side
    "```{r tab_{{j}}_OSL_plots}\n", 
    "record_1 <- data_set[[{{j}}]]@records[[record_index[[{{j}}]][1]]]\n",
    "plot_1 <- plot_OSLcurve(record_1, record_1@info$COMPONENTS,
                             display = \"compare_lin\", hide_plot = TRUE)\n",
    "record_2 <- data_set[[{{j}}]]@records[[record_index[[{{j}}]][2]]]\n",
    "plot_2 <- plot_OSLcurve(record_2, record_2@info$COMPONENTS,
                             display = \"compare_lin\", hide_plot = TRUE)\n",
    "grid.arrange(plot_1, plot_2, layout_matrix = t(matrix(c(1,2))))\n",
    "```\n",
    "<i><b>Figure 2.`r {{j}}`:</b> Signal components and residual curve of first two `r record_type` records.<br>",
    "<u>Left:</u> First record (index = `r record_index[[{{j}}]][1]`),",
    " usually natural or recovered dose signal ($L_n$).<br>",
    "<u>Right:</u> Second record (index = `r record_index[[{{j}}]][2]`),",
    " usually test-dose signal ($T_n$).</i><br><br>\n")

### ALTERNATIVE MORE VARIABLE (but buggy) PLOT ASSIGNEMENT:
#    "plot_objects <- list()\n",
#    "for (i in record_index[[{{j}}]][1:2]) {\n",
#    "record <- data_set[[{{j}}]]@records[[i]]\n",
#    "plot_objects <- c(plot_objects, 
#                      plot_OSLcurve(record,
#                                     record@info$COMPONENTS,
#                                     display = \"compare_lin\",
#                                     hide_plot = TRUE))}\n",

tabs <- lapply(as.list(1:length(data_set)),
                function(j) knitr::knit_expand(text = dynamic_code))
```
`r knitr::knit(text = unlist(tabs))`


## Further statistics

<br><br><i>

<u>Planned features:</u>  

* Component wise barplot and histogram of natural signal intensities
* Initial signal composition of first and second measurement
* Scatterplots or full correlogram of the components and the irradiation time


```{r some_infos}
# Hint: search for "Correlogram"
# or: http://www.sthda.com/english/wiki/correlation-analyses-in-r

#pairs(subset(results, select = c("n.1", "n.2", "n.3", "IRR_TIME")))

```

</i><br><br>

***
<br><br>

## Algorithm description

In Step 2, we decompose each OSL curve into its signal components. We set the decay constants found in Step 1 as fixed values for all OSL curves of the data set. 

### Signal bin intervals

### Decomposition

-------------------

1. Divide the measurement time into *K* intervals. These intervals are calculated and optimized globally by `calc_OSLintervals()`.
2. Integrate the signal curve of each OSL record over these intervals. From the integration values and the fitting model found in Step 1, build one equation system with *K* equations for each OSL record. 
3. Solve the equation system by an analytic determinant based method, called 'Cramer's rule', and get the area under the component curve or 'intensity' *n~k~* for each signal component
4. To enhance stability and precision of the method, refine the set of *n~k~* values in a quasi-linear regression using `base::nls()`. If this refining-fit fails, go on with the Cramer's rule achieved values.
5. Calculate the standard deviation of the integration values from step 2 by the residuals between fit-model OSL curve and real data points 
6. Apply the propagation of uncertainty method onto Cramer's rule and calculate the uncertainty $\sigma_k$ for each component intensity value *n~k~*

-------------------

### Result refinement


## Further information

### Disclaimer

This report was automatically generated by functions of the **R** package `OSLdecomposition` written and maintained by Dirk Mittelstraß (<dirk.mittelstrass@luminescence.de>).

This report and the containing results can be used, shared and published by the data set maintainer at will. If the results are published, however, it is demanded to state the main **R** package `OSLdecomposition` including its version number (`r packageVersion("OSLdecomposition")`). It is also recommended to add this report to the supplement of your publication.   

A full description of the method and the algorithms involved, as well as some performance tests, can be found in the master thesis this script is based upon:

-------------------

*D. Mittelstraß, ‘Decomposition of weak optically stimulated luminescence signals and its application in retrospective dosimetry at quartz’, Master thesis, TU Dresden, Dresden, 2019.*

-------------------


### Script settings


**Algorithm parameters** | 
-----------------------|----------------
Applied algorithm | `r dec_data$parameters$algorithm`
Error estimation approach | `r dec_data$parameters$error_calculation`
With background component? | `r dec_data$parameters$background_fitting`

### Session info

```{r session_info}

print(sessionInfo(), locale = FALSE)
```

## References





```{r curve_viewer, eval=FALSE, include=FALSE}

#### CURVE VIEWER ####
# does not work with static Rmarkdown documents

# Determine zoom minimum (10 data points)
#zoom_min <- ceiling(10*100 / length(global_curve$time)) / 100

# Create shiny plot to surf through curves
numericInput("aliquot_i", "Aliquot no.: ", value = 1, min = 1, max = length(data_set))
numericInput("record_i", "Record no.: ", value = 1, min = 1, max = length(record_index))

#sliderInput("aliquot_i", "Aliquot no.: ", value = 1, step = 1, min = 1, max = aliquot.n)
#sliderInput("record_i", "Curve no.: ", value = 1, step = 1, min = 1, max = length(record_index))
sliderInput("zoom", "X axis zoom", min = 0.1, max = 1, value = 1)


renderPlot(width = 900, height = 450,res = 100,
   {
     i <- record_index[input$record_i]
     record <- data_set[[input$aliquot_i]]@records[[i]]
     plot_title <- paste0("Record ", i, " of aliquot ", input$aliquot_i,
                          " (pos.=", record@info[["POSITION"]], "): ",
                          record@info[["DTYPE"]], 
                          " (dose=", record@info[["IRR_TIME"]], ")")
    
     plot_OSLcurve(record, 
                  record@info[["COMPONENTS"]], 
                  display = "detailed", 
                  title = plot_title,
                  algorithm = dec_data$parameters$algorithm,
                  zoom = input$zoom)
})

```

